Jaigo Technical Report

### 2007-12-19 trunk/paper/paper.tex
editing for style and punctuation

While chess eventually yielded to clever hardware and algorithms, the sheer state-space of Go has rendered specialize search approaches weak. In 1965, noted statistician I.J. Good noted presciently:

\begin{quote}
Go on a computer? – In order to programme [sic] a computer to play a reasonable game of Go – rather than merely a legal game – it is necessary to formalise the principles of good strategy, or to design a learning programme. The principles are more qualitative and mysterious than in chess, and depend more on judgment. So I think it will be even more difficult to programme a computer to play a reasonable game of Go than of chess. \cite{Good65}
\end{quote}

@@

The exploration of alternative future board states often yields duplicate (potential) states; it is helpful to memorize board evaluation such that a board state may be evaluated only once, regardless of the paths by which it is reached \cite{Zobrist1970b}. The Zobrist hash function is useful in this capacity: (reduce xor (map f board-positions)), f: (x, y, color) -> Z.

@@

Third, while GA \& NN initially seemed promising, no commercial package uses them due to lack of speed.

Thus, if all older techniques are inadequate and new techniques do not pan out, whither computer Go? Our project seeks to create a JavaScript-based Go-engine. JavaScript is a dynamically typed, imperative language that supports the functional and object-oriented paradigms \cite{ECMA-262}. It is ubiquitous due to its inclusion in nearly every Web browser, and recently has enjoyed a surge of popularity among programmers due to the availability of cross-platform libraries and the success of applications such as GMail. As professional programmers, we viewed this project as an opportunity to explore not only AI but also renewed interest in the JavaScript language. Also, JavaScript is the only supported means of developing iPhone applications. iPhone is an intrinsically interesting platform; its communication capabilities and computational limitations only heighten the interest.

@@

We chose to implement two different move choice engines, each of which provides a different playing experience. SimpleGo v1, a 50kyu greedy search move generator which picks a random move from list of moves with worst opponent score. Crawler, an unranked move generator which analyzes the current blocks on the board and picks the move that both connects to a block and results in the largest number of liberties for the block.

\chapter{Technical details}
Code Listing
- Implementation of Benson's algorithm for determining unconditional life: Analyze\_Color\_Unconditional\_Status
- Greedy Search algorithm: select\_scored\_move
- Crawler algorithm: select\_crawler\_move

\chapter{Discussions, observations, and comparisons}
We implemented two move choice engines based on two different algorithms, scoring for color and blocks based on Benson's algorithm, and representations of the board, the game, eyes, blocks, and numerous language-specific utilities. Our goals were compactness and speed due to the target platform limitations.

@@

\chapter{Recommendations, especially for future work and unsolved issues}
We feel that the choice of browser is more important than the choice of computing platform for the performance of jaigo. However, we have not quantified this claim.

>>>>

While chess eventually yielded to clever hardware and algorithms, the sheer state-space of Go has rendered specialize search approaches weak. In 1965, noted statistician I.J. Good presciently commented:

\begin{quote}
Go on a computer? -- In order to programme [sic] a computer to play a reasonable game of Go -- rather than merely a legal game -- it is necessary to formalise the principles of good strategy, or to design a learning programme. The principles are more qualitative and mysterious than in chess, and depend more on judgment. So I think it will be even more difficult to programme a computer to play a reasonable game of Go than of chess. \cite{Good65}
\end{quote}

@@

The exploration of alternative future board states often yields duplicate (potential) states; it is helpful to memorize board evaluation such that a board state may be evaluated only once, regardless of the paths by which it is reached \cite{Zobrist1970b}. The Zobrist hash function is useful in this capacity: (reduce xor (map f board-positions)), $f: (x, y, color) \rightarrow Z$.

@@

Third, while genetic algorithms and neural networks initially seemed promising, no commercial package uses them due to lack of speed.

Thus, if all older techniques are inadequate and new techniques do not pan out, whither computer Go? Our project seeks to create a JavaScript-based Go-engine. JavaScript is a dynamically typed, imperative language that supports the functional and object-oriented paradigms \cite{ECMA-262}. It is ubiquitous due to its inclusion in nearly every Web browser, and recently has enjoyed a surge of popularity among programmers due to the availability of cross-platform libraries and the success of applications such as GMail. As professional programmers, we viewed this project as an opportunity to explore not only AI but also to pursue our renewed interest in the JavaScript language. Also, JavaScript is the only supported means of developing iPhone applications. iPhone is an intrinsically interesting platform; its communication capabilities and computational limitations heighten the interest.

@@

We implemented two different move choice engines, each of which provides a different playing experience. SimpleGo v1, a 50kyu greedy search move generator which picks a random move from list of moves with worst opponent score. Crawler, an unranked move generator which analyzes the current blocks on the board and picks the move that both connects to a block and results in the largest number of liberties for the block.

\chapter{Technical details}
%Code Listing
- Implementation of Benson's algorithm for determining unconditional life: Analyze\_Color\_Unconditional\_Status
- Greedy Search algorithm: select\_scored\_move
- Crawler algorithm: select\_crawler\_move

\chapter{Discussions, observations, and comparisons}
We implemented representations of the board, the game, eyes, and blocks; scoring for color and blocks based on Benson's algorithm; two move choice engines based on two different algorithms; and numerous language-specific utilities. Our goals were compactness and speed due to the target platform limitations.

@@

\chapter{Recommendations, especially for future work and unsolved issues}
We feel that the choice of browser is more important than the choice of computing platform for the performance of jaigo. Our anecdotal experience and those of others on the Web show that there is an order of magnitude difference between popular browsers for some JavaScript operations. However, we have not carefully studied this issue.%http://therealcrisp.xs4all.nl/blog/2006/12/09/string-performance-in-internet-explorer/
